% TODO: figures and chapters need independent label indexing

\documentclass{report}
\usepackage{array}
\usepackage{listings}
\usepackage{graphics}
\usepackage{setspace}
\usepackage{amsmath}

\setlength{\textwidth}{6.5in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}

\newlength{\drop}

\newcommand*{\titleHTMmath}{\begingroup%
\drop=0.08\textheight
\vspace*{\drop}
\begin{center}
{\Large Mathematics of\par}
{\Large Hierarchical Temporal Memory\par}
\vfill
{\Large Version 1.0, 14th April, 2015\par}
{\Large \textcopyright{} Numenta, Inc. 2011\par}
{\Large \textcopyright{} Numenta and the NuPIC Community 2014\par}
\end{center}
\vfill
\endgroup}


\begin{document}
\lstset{language=Python}        % TODO: good choice for pseudocode


% TODO: Numenta logo

\pagestyle{empty}
\titleHTMCLA

Use of Numenta's software and intellectual property, including the
ideas contained in this document, are free for non-commercial research
purposes. For details, see
http://www.numenta.com/about-numenta/licensing.php.

\clearpage

\subsection*{Read This First!}

This is a draft version of this document. There are several things
missing that you should be aware of.

\subsection*{What IS in this document:}
This document describes in detail new algorithms for learning and
prediction developed by Numenta. The new algorithms are described in
sufficient detail that a programmer can understand and implement them
if desired. It starts with an introductory chapter. If you have been
following Numenta and have read some of our past white papers, the
material in the introductory chapter will be familiar. The other
material is new.

There are several topics related to the implementation of these new
algorithms that did not make it into this early draft.

\subsection*{What is NOT in this document:}

\begin{itemize}

\item Although most aspects of the algorithms have been implemented
and tested in software, none of the test results are currently
included.

\item The algorithms are capable of on-line learning. A few details
needed to fully implement on-line learning in some rarer cases are not
described.

\end{itemize}

We are making this document available in its current form because we
think the algorithms will be of interest to others. The missing
components of the document should not impede understanding and
experimenting with the algorithms by motivated researchers. We will
revise this document regularly to reflect our progress.

\tableofcontents

% Table of Contents
% Preface 4
% Chapter 1: HTM Overview 7
% Chapter 2: HTM Cortical Learning Algorithms 19
% Chapter 3: Spatial Pooling Implementation and Pseudocode 34
% Chapter 4: Temporal Pooling Implementation and Pseudocode 39
% Appendix A: A Comparison between Biological Neurons 47 and HTM Cells
% Appendix B: A Comparison of Layers in the Neocortex and 54 an HTM Region
% Glossary 65

\input{010-preface.tex}
\input{020-htm-overview.tex}
\input{030-htm-cla.tex}
\input{040-recognition-implementation.tex}
\input{050-prediction-implementation.tex}
\input{055-sensorimotor-cla.tex}

\appendix

\input{060-neurons-vs-htm-cells.tex}
\input{070-layers-vs-htm-regions.tex}
\input{080-glossary.tex}


\end{document}

